---
title: "10x_run1_DoubletFinder"
author: "SandraTietscher"
date: "2019 M06 17"
output: html_document
---


## Load libraries and data

```{r}
if (!require(ggplot2)) install.packages('ggplot2')
library(ggplot2)

if (!require(dplyr)) install.packages('dplyr')
library(dplyr)

if (!require(Seurat)) install.packages('Seurat')
library(Seurat)

#if (!require(DoubletFinder)) #devtools::install_github('chris-mcginnis-ucsf/DoubletFinder')
library(DoubletFinder)


in.path = "C:/Users/SandraT/BBlab/Macrophages/Sorting/scRNA_seq/Data/10x/Run_02/CellRangerCount_40342_2019-10-15--13-52-11/"
out.path = "Z:/server_homes/sandrat/Code/10x/DoubletFinder/output/TBB184/"
TBB184.data <- Read10X(paste0(in.path, "TBB184/filtered_feature_bc_matrix"))
```

# Individual Sample preprocessing

```{r}

# Create Seurat object
TBB184 <- CreateSeuratObject(counts = TBB184.data, min.cells = 5, min.features = 200, project = "TBB184")
TBB184 <- AddMetaData(TBB184, "TIG2", col.name = 'TIG')
TBB184 <- AddMetaData(TBB184, "B5", col.name = 'batch')
rm(TBB184.data)

# Calculate mitotic percentage
mito.features <- grep(pattern = "^MT-", x = rownames(x = TBB184), value = TRUE)
percent.mito <- Matrix::colSums(x = GetAssayData(object = TBB184, slot = 'counts')[mito.features, ]) / Matrix::colSums(x = GetAssayData(object = TBB184, slot = 'counts'))
TBB184[['percent.mito']] <- percent.mito

# Pre-process Seurat object
TBB184 <- NormalizeData(TBB184)
TBB184 <- ScaleData(TBB184)
TBB184 <- FindVariableFeatures(TBB184, x.low.cutoff = 0.0125, y.cutoff = 0.25, do.plot=FALSE)
TBB184 <- RunPCA(TBB184, features = VariableFeatures(object = TBB184), verbose = FALSE)
#TBB184 <- RunTSNE(object = TBB184, dims = 1:10)
TBB184 <- RunUMAP(object = TBB184, dims = 1:10)

FeaturePlot(TBB184, reduction = 'umap', features = "nFeature_RNA")
DimPlot(TBB184, label = TRUE, group.by = "RNA_snn_res.0.4")
DimPlot(TBB184, label = TRUE, reduction = 'umap')
```

## Graph-based clustering
Low resolution clustering as basis for homotypic doublet model.

```{r}
TBB184 <- FindNeighbors(object = TBB184, dims = 1:10)
TBB184 <- FindClusters(object = TBB184, resolution = 0.4)

VlnPlot(object = TBB184, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3, pt.size = 0, sort = FALSE)
FeaturePlot(TBB184, reduction = 'umap', features = c("CD3E", "CD14", "PTPRC", "EPCAM", "PECAM1", "FAP"))
```
## Assigning cell type identity to clusters

```{r}
#Read in table with cluster ids
cluster.ids <- read.csv(file = paste0(out.path, "cluster_celltype.csv"))

new.cluster.ids <- cluster.ids$celltype
new.cluster.ids <- as.character(new.cluster.ids)
names(x = new.cluster.ids) <- levels(x = TBB184)
TBB184 <- RenameIdents(object = TBB184, new.cluster.ids)

#Save celltype information in metadata
TBB184[["annotation"]] <- Idents(object = TBB184)

# plot_celltype 
DimPlot(object = TBB184, reduction = 'umap', label = FALSE, pt.size = 0.5, group.by = "annotation")
```
## pK identification

```{r}

sweep.res.list_TBB184 <- paramSweep_v3(TBB184, PCs = 1:10)
sweep.stats_TBB184 <- summarizeSweep(sweep.res.list_TBB184, GT = FALSE)
bcmvn_TBB184 <- find.pK(sweep.stats_TBB184)
write.csv(bcmvn_TBB184, file=paste0(out.path, "pK_identification.csv"))

```

## Homotypic Doublet Proportion Estimate 
```{r}
annotations <- TBB184@meta.data$annotation
homotypic.prop <- modelHomotypic(annotations)           ## ex: annotations ~ number of cell types present
nExp_poi <- round(0.105*length(TBB184@meta.data$nFeature_RNA))  ## Assuming 8% doublet formation rate - tailor for your dataset
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
```


## Run DoubletFinder with varying classification stringencies 
```{r}
TBB184 <- doubletFinder_v3(TBB184, PCs = 1:10, pN = 0.25, pK = 0.02, nExp = nExp_poi, reuse.pANN = FALSE)
TBB184 <- doubletFinder_v3(TBB184, PCs = 1:10, pN = 0.25, pK = 0.02, nExp = nExp_poi.adj, reuse.pANN = "pANN_0.25_0.02_1366")
```

## Plot results 
```{r}
TBB184@meta.data[,"DF_hi.lo"] <- TBB184@meta.data$DF.classifications_0.25_0.02_1366
TBB184@meta.data$DF_hi.lo[which(TBB184@meta.data$DF_hi.lo == "Doublet" & TBB184@meta.data$DF.classifications_0.25_0.02_1054 == "Singlet")] <- "Doublet_lo"
TBB184@meta.data$DF_hi.lo[which(TBB184@meta.data$DF_hi.lo == "Doublet")] <- "Doublet_hi"
DFplot_105perc_pK0.13 <- DimPlot(TBB184, group.by="DF_hi.lo", plot.order=c("Doublet_hi","Doublet_lo","Singlet"), colors.use=c("black","gold","red"))

ggsave(paste0(out.path, "umap_doublets_pk0.02.png"))
```

## Sanity checks

```{r}

# find all markers of cluster 15
#cluster15.markers <- FindMarkers(object = TBB184, ident.1 = 15, min.pct = 0.25)
#head(x = cluster15.markers, n = 15)

DimPlot(TBB184, label = TRUE)
DotPlot(object = TBB184, features = c("CD3E", "CD14", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB184, features = c("CD3E", "CD14", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB184, features = c("FAP", "PECAM1", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB184, features = c("CD14", "MKI67"))
FeaturePlot(object = TBB184, features = c("MKI67"), split.by = "DF_hi.lo")

FeatureScatter(TBB184, feature1 = "CD3E", feature2 = "MS4A1", group.by = "DF_hi.lo")
ggsave(paste0(out.path, "FeatureScatter_pk0.02_CD3_CD20.png"))

table(TBB184@meta.data$DF_hi.lo, TBB184@meta.data$annotation)
```

## Create a subset excluding high-confidence doublets
```{r}
Idents(TBB184) <- TBB184@meta.data$DF_hi.lo

# Create a Seurat object with only the immune cell clusters
TBB184.singlet <- subset(x = TBB184, idents =c("Singlet", "Doublet_lo"))

#Test
DimPlot(object = TBB184.singlet, reduction = 'umap', pt.size = 0.5)

#Remove unnecessary Metadata columns
TBB184.singlet$RNA_snn_res.0.4 <- NULL
TBB184.singlet$seurat_clusters <- NULL
TBB184.singlet$annotation <- NULL
TBB184.singlet$pANN_0.25_0.16_1366 <- NULL
TBB184.singlet$pANN_0.25_0.13_1366 <- NULL
TBB184.singlet$pANN_0.25_0.02_1366 <- NULL
TBB184.singlet$DF.classifications_0.25_0.16_1366 <- NULL
TBB184.singlet$DF.classifications_0.25_0.16_1054 <- NULL
TBB184.singlet$DF.classifications_0.25_0.13_1366 <- NULL
TBB184.singlet$DF.classifications_0.25_0.13_1054 <- NULL
TBB184.singlet$DF.classifications_0.25_0.02_1366 <- NULL
TBB184.singlet$DF.classifications_0.25_0.02_1054 <- NULL
#TBB184.singlet$DF_hi.lo<- NULL

#Save Seurat object
saveRDS(TBB184.singlet, file = paste0(out.path, "TBB184_singlet.rds"))
```