---
title: "10x_run1_DoubletFinder"
author: "SandraTietscher"
date: "2019 M10 25"
output: html_document
---


## Load libraries and data

```{r}
if (!require(ggplot2)) install.packages('ggplot2')
library(ggplot2)

if (!require(dplyr)) install.packages('dplyr')
library(dplyr)

if (!require(Seurat)) install.packages('Seurat')
library(Seurat)

#if (!require(DoubletFinder)) devtools::install_github('chris-mcginnis-ucsf/DoubletFinder')
library(DoubletFinder)


in.path = "~/Data/Sequencing_data/10x_run2_cellranger_FilteredFeatureMatrix/"
out.path = "~/Code/10x/DoubletFinder/output/TBB035/"
TBB035.data <- Read10X(paste0(in.path, "TBB035"))
```

# Individual Sample preprocessing

```{r}

# Create Seurat object
TBB035 <- CreateSeuratObject(counts = TBB035.data, min.cells = 5, min.features = 200, project = "TBB035")
TBB035 <- AddMetaData(TBB035, "TIG3", col.name = 'TIG')
TBB035 <- AddMetaData(TBB035, "B5", col.name = 'batch')
rm(TBB035.data)

# Calculate mitotic percentage
mito.features <- grep(pattern = "^MT-", x = rownames(x = TBB035), value = TRUE)
percent.mito <- Matrix::colSums(x = GetAssayData(object = TBB035, slot = 'counts')[mito.features, ]) / Matrix::colSums(x = GetAssayData(object = TBB035, slot = 'counts'))
TBB035[['percent.mito']] <- percent.mito

# Pre-process Seurat object
TBB035 <- NormalizeData(TBB035)
TBB035 <- ScaleData(TBB035)
TBB035 <- FindVariableFeatures(TBB035, x.low.cutoff = 0.0125, y.cutoff = 0.25, do.plot=FALSE)
TBB035 <- RunPCA(TBB035, features = VariableFeatures(object = TBB035), verbose = FALSE)
#TBB035 <- RunTSNE(object = TBB035, dims = 1:10)
TBB035 <- RunUMAP(object = TBB035, dims = 1:10)

FeaturePlot(TBB035, reduction = 'umap', features = "nFeature_RNA")
#DimPlot(TBB035, label = TRUE, group.by = "RNA_snn_res.0.4")
DimPlot(TBB035, label = TRUE, reduction = 'umap')
```

## Graph-based clustering
Low resolution clustering as basis for homotypic doublet model.

```{r}
TBB035 <- FindNeighbors(object = TBB035, dims = 1:10)
TBB035 <- FindClusters(object = TBB035, resolution = 0.4)

VlnPlot(object = TBB035, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3, pt.size = 0, sort = FALSE)
FeaturePlot(TBB035, reduction = 'umap', features = c("CD3E", "CD14", "PTPRC", "EPCAM", "PECAM1", "FAP"))
```
## Assigning cell type identity to clusters

```{r}
#Read in table with cluster ids
cluster.ids <- read.csv(file = paste0(out.path, "cluster_celltype.csv"))

new.cluster.ids <- cluster.ids$celltype
new.cluster.ids <- as.character(new.cluster.ids)
names(x = new.cluster.ids) <- levels(x = TBB035)
TBB035 <- RenameIdents(object = TBB035, new.cluster.ids)

#Save celltype information in metadata
TBB035[["annotation"]] <- Idents(object = TBB035)

# plot_celltype 
DimPlot(object = TBB035, reduction = 'umap', label = FALSE, pt.size = 0.5, group.by = "annotation")
```
## pK identification

```{r}

sweep.res.list_TBB035 <- paramSweep_v3(TBB035, PCs = 1:10)
sweep.stats_TBB035 <- summarizeSweep(sweep.res.list_TBB035, GT = FALSE)
bcmvn_TBB035 <- find.pK(sweep.stats_TBB035)
write.csv(bcmvn_TBB035, file=paste0(out.path, "pK_identification.csv"))

```

## Homotypic Doublet Proportion Estimate 
```{r}
annotations <- TBB035@meta.data$annotation
homotypic.prop <- modelHomotypic(annotations)           ## ex: annotations ~ number of cell types present
nExp_poi <- round(0.08*length(TBB035@meta.data$nFeature_RNA))  ## Assuming 8% doublet formation rate - tailor for your dataset
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
```


## Run DoubletFinder with varying classification stringencies 
```{r}
TBB035 <- doubletFinder_v3(TBB035, PCs = 1:10, pN = 0.25, pK = 0.005, nExp = nExp_poi, reuse.pANN = FALSE)
TBB035 <- doubletFinder_v3(TBB035, PCs = 1:10, pN = 0.25, pK = 0.005, nExp = nExp_poi.adj, reuse.pANN = "pANN_0.25_0.005_910")
```

## Plot results 
```{r}
TBB035@meta.data[,"DF_hi.lo"] <- TBB035@meta.data$DF.classifications_0.25_0.005_910
TBB035@meta.data$DF_hi.lo[which(TBB035@meta.data$DF_hi.lo == "Doublet" & TBB035@meta.data$DF.classifications_0.25_0.005_707 == "Singlet")] <- "Doublet_lo"
TBB035@meta.data$DF_hi.lo[which(TBB035@meta.data$DF_hi.lo == "Doublet")] <- "Doublet_hi"
DFplot_7perc_pK0.005 <- DimPlot(TBB035, group.by="DF_hi.lo", plot.order=c("Doublet_hi","Doublet_lo","Singlet"), colors.use=c("black","gold","red"))

ggsave(paste0(out.path, "umap_doublets_pk0.005.png"))
```

## Sanity checks

```{r}

# find all markers of cluster 15
#cluster15.markers <- FindMarkers(object = TBB035, ident.1 = 15, min.pct = 0.25)
#head(x = cluster15.markers, n = 15)

DimPlot(TBB035, label = TRUE)
DotPlot(object = TBB035, features = c("CD3E", "CD14", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB035, features = c("CD3E", "CD14", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB035, features = c("FAP", "PECAM1", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB035, features = c("CD14", "MKI67"))
FeaturePlot(object = TBB035, features = c("MKI67"), split.by = "DF_hi.lo")

FeatureScatter(TBB035, feature1 = "CD14", feature2 = "EPCAM", group.by = "DF_hi.lo")
ggsave(paste0(out.path, "FeatureScatter_pk0.005_CD14_Epcam.png"))

table(TBB035@meta.data$DF_hi.lo, TBB035@meta.data$annotation)
```

## Create a subset excluding high-confidence doublets
```{r}
Idents(TBB035) <- TBB035@meta.data$DF_hi.lo

# Create a Seurat object with only the immune cell clusters
TBB035.singlet <- subset(x = TBB035, idents =c("Singlet", "Doublet_lo"))

#Test
DimPlot(object = TBB035.singlet, reduction = 'umap', pt.size = 0.5)

#Remove unnecessary Metadata columns
TBB035.singlet$RNA_snn_res.0.4 <- NULL
TBB035.singlet$seurat_clusters <- NULL
TBB035.singlet$annotation <- NULL
TBB035.singlet$pANN_0.25_0.24_910 <- NULL
TBB035.singlet$pANN_0.25_0.005_910 <- NULL
TBB035.singlet$DF.classifications_0.25_0.24_707 <- NULL
TBB035.singlet$DF.classifications_0.25_0.24_910 <- NULL
TBB035.singlet$DF.classifications_0.25_0.005_910 <- NULL
TBB035.singlet$DF.classifications_0.25_0.005_707 <- NULL
#TBB035.singlet$DF_hi.lo<- NULL

#Save Seurat object
saveRDS(TBB035.singlet, file = paste0(out.path, "TBB035_singlet.rds"))
