---
title: "10x_run1_DoubletFinder"
author: "SandraTietscher"
date: "2019 M06 17"
output: html_document
---


## Load libraries and data

```{r}
if (!require(ggplot2)) install.packages('ggplot2')
library(ggplot2)

if (!require(dplyr)) install.packages('dplyr')
library(dplyr)

if (!require(Seurat)) install.packages('Seurat')
library(Seurat)

if (!require(DoubletFinder)) devtools::install_github('chris-mcginnis-ucsf/DoubletFinder')
library(DoubletFinder)

#if (!require(reticulate)) install.packages("reticulate")
#library(reticulate)
#use_condaenv("r-reticulate")
#conda_install("r-reticulate", "umap-learn")
#reticulate::py_install(packages ='umap-learn')
#use_python("C:/Users/SandraT/AppData/Local/Continuum/anaconda3/envs/r-reticulate/python.exe")
#umap-learn <- import("umap-learn")

in.path = "~/RNAseq/fgcz/ST_TAMsBC_10x_Run01/CellRangerCount_37250_2019-05-27--16-31-27/"
out.path = "~/Code/10x/DoubletFinder/output/TBB129/"
TBB129.data <- Read10X(paste0(in.path, "T_BB129/filtered_feature_bc_matrix"))
#TBB214.data <- Read10X(paste0(in.path, "T_BB214/filtered_feature_bc_matrix"))
```

# Individual Sample preprocessing

```{r}

# Create Seurat object
TBB129 <- CreateSeuratObject(counts = TBB129.data, min.cells = 5, min.features = 200, project = "TBB129")
TBB129 <- AddMetaData(TBB129, "TIG2", col.name = 'TIG')
TBB129 <- AddMetaData(TBB129, "B1", col.name = 'batch')
rm(TBB129.data)

# Calculate mitotic percentage
mito.features <- grep(pattern = "^MT-", x = rownames(x = TBB129), value = TRUE)
percent.mito <- Matrix::colSums(x = GetAssayData(object = TBB129, slot = 'counts')[mito.features, ]) / Matrix::colSums(x = GetAssayData(object = TBB129, slot = 'counts'))
TBB129[['percent.mito']] <- percent.mito

# Pre-process Seurat object
TBB129 <- NormalizeData(TBB129)
TBB129 <- ScaleData(TBB129)
TBB129 <- FindVariableFeatures(TBB129, x.low.cutoff = 0.0125, y.cutoff = 0.25, do.plot=FALSE)
TBB129 <- RunPCA(TBB129, features = VariableFeatures(object = TBB129), verbose = FALSE)
#TBB129 <- RunTSNE(object = TBB129, dims = 1:10)
TBB129 <- RunUMAP(object = TBB129, dims = 1:10)

#FeaturePlot(TBB129, reduction = 'tsne', features = "nFeature_RNA")
#DimPlot(TBB129, label = TRUE)
```

## Graph-based clustering
Low resolution clustering as basis for homotypic doublet model.

```{r}
TBB129 <- FindNeighbors(object = TBB129, dims = 1:10)
TBB129 <- FindClusters(object = TBB129, resolution = 0.4)

VlnPlot(object = TBB129, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3, pt.size = 0, sort = FALSE)
```
## Assigning cell type identity to clusters

```{r}
#Read in table with cluster ids
cluster.ids <- read.csv(file = paste0(out.path, "cluster_celltype.csv"))

new.cluster.ids <- cluster.ids$celltype
new.cluster.ids <- as.character(new.cluster.ids)
names(x = new.cluster.ids) <- levels(x = TBB129)
TBB129 <- RenameIdents(object = TBB129, new.cluster.ids)

#Save celltype information in metadata
TBB129[["annotation"]] <- Idents(object = TBB129)

# plot_celltype 
#DimPlot(object = TBB129, reduction = 'tsne', label = FALSE, pt.size = 0.5, group.by = "RNA_snn_res.0.4")
```
## pK identification

```{r}

#sweep.res.list_TBB129 <- paramSweep_v3(TBB129, PCs = 1:10)
#sweep.stats_TBB129 <- summarizeSweep(sweep.res.list_TBB129, GT = FALSE)
#bcmvn_TBB129 <- find.pK(sweep.stats_TBB129)

```

## Homotypic Doublet Proportion Estimate 
```{r}
annotations <- TBB129@meta.data$annotation
homotypic.prop <- modelHomotypic(annotations)           ## ex: annotations ~ number of cell types present
nExp_poi <- round(0.12*length(TBB129@meta.data$nFeature_RNA))  ## Assuming 8% doublet formation rate - tailor for your dataset
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
```


## Run DoubletFinder with varying classification stringencies 
```{r}
TBB129 <- doubletFinder_v3(TBB129, PCs = 1:10, pN = 0.25, pK = 0.005, nExp = nExp_poi, reuse.pANN = FALSE)
TBB129 <- doubletFinder_v3(TBB129, PCs = 1:10, pN = 0.25, pK = 0.005, nExp = nExp_poi.adj, reuse.pANN = "pANN_0.25_0.005_2275")
```

## Plot results 
```{r}
TBB129@meta.data[,"DF_hi.lo"] <- TBB129@meta.data$DF.classifications_0.25_0.005_2275
TBB129@meta.data$DF_hi.lo[which(TBB129@meta.data$DF_hi.lo == "Doublet" & TBB129@meta.data$DF.classifications_0.25_0.005_1644 == "Singlet")] <- "Doublet_lo"
TBB129@meta.data$DF_hi.lo[which(TBB129@meta.data$DF_hi.lo == "Doublet")] <- "Doublet_hi"
DFplot_12perc_pK0.005 <- UMAPPlot(TBB129, group.by="DF_hi.lo", plot.order=c("Doublet_hi","Doublet_lo","Singlet"), colors.use=c("black","gold","red"))
```

## Sanity checks

```{r}

# find all markers of cluster 15
#cluster15.markers <- FindMarkers(object = TBB129, ident.1 = 15, min.pct = 0.25)
#head(x = cluster15.markers, n = 15)

DimPlot(TBB129, label = TRUE)
DotPlot(object = TBB129, features = c("CD3E", "CD14", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB129, features = c("CD3E", "CD14", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB129, features = c("FAP", "PECAM1", "PTPRC", "EPCAM"))
FeaturePlot(object = TBB129, features = c("CD14", "MKI67"))
FeaturePlot(object = TBB129, features = c("MKI67"), split.by = "DF_hi.lo")

table(TBB129@meta.data$DF_hi.lo, TBB129@meta.data$RNA_snn_res.0.4)
```

## Create a subset excluding high-confidence doublets
```{r}
Idents(TBB129) <- TBB129@meta.data$DF_hi.lo

# Create a Seurat object with only the immune cell clusters
TBB129.singlet <- subset(x = TBB129, idents = c("Singlet", "Doublet_lo"))

#Remove unnecessary Metadata columns
TBB129.singlet$RNA_snn_res.0.4 <- NULL
TBB129.singlet$seurat_clusters <- NULL
TBB129.singlet$annotation <- NULL
TBB129.singlet$DF_hi.lo<- NULL
TBB129.singlet$pANN_0.25_0.005_2275<- NULL
TBB129.singlet$DF.classifications_0.25_0.005_2275<- NULL
TBB129.singlet$DF.classifications_0.25_0.005_1644<- NULL


#Test
DimPlot(object = TBB129.singlet, reduction = 'tsne', pt.size = 0.5)

#Save Seurat object
saveRDS(TBB129.singlet, file = paste0(out.path, "TBB129_singlet.rds"))
```