---
title: "Fig4_signatures"
author: "Sandra Tietscher"
date: "5/5/2020"
output: html_document
---

## Load libraries and data

```{r}
library(ggplot2)
library(Seurat)
library(tidyverse)
library(data.table)
library(magrittr)
library(viridis)
library(patchwork)
library(ggrepel)
library(ggpubr)

in.path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/myeloid/"
out.path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/figures_for_paper/Fig4/" 
all.myeloid <- readRDS(file= paste0(in.path, "myeloid_excl_cl9.rds"))
```

## Define signature genes

```{r}
T_attr <- c("CCL21", "CCL17",  "CCL2", "CXCL9", "CXCL10", "CXCL11", "CXCL12", "CCL3", "CCL4", "CCL5", "CXCL16")

T_supp <- c("CD80", "CD86", "IDO1", "CD274", "PDCD1LG2",  "TNFSF10", "HLA-E", "HLA-G", "VTCN1", "IL10", "TGFB1", "TGFB2", "PTGS2", "PTGES", "LGALS9", "CCL22")

M1_features <- c("NOS2", "IL12A", "IL12B", "FCGR1A", "FCGR1B", "CD80", "IL23A", "CXCL9", "CXCL10", "CD86", "IL1A", "IL1B", "IL6", "TNF", "CCL5", "IRF5", "IRF1", "CD40", "IDO1", "KYNU", "CCR7" )

M2_features <- c("ARG2", "IL10", "FCGR2A", "FCGR2B", "FCGR2C", "CD200R1", "PDCD1LG2", "CD274", "MARCO", "CSF1R", "MRC1", "IL1RN", "IL1R2", "IL4R", "CCL4", "CCL13", "CCL20", "CCL17", "CCL18", "CCL22", "LYVE1", "VEGFA", "VEGFB", "VEGFC", "VEGFD", "EGF", "CTSA", "CTSB", "CTSC", "CTSD","TGFB1", "TGFB2", "TGFB3", "MMP14", "MMP19", "MMP9", "CLEC7A", "WNT7B", "FASLG", "TNFSF12", "TNFSF8", "CD276", "VTCN1", "MSR1", "FN1", "IRF4")

# Signature of 15 random genes (as a control)
set.seed(20201130)
random <- sample(rownames(all.myeloid), 15)
```

### Calculate T-cell-associated signatures using log-transformation of percentages for each gene
To avoid dominance of few genes when calculating the signature score: Log-normalization

1. For each cell, divide counts of gene i through total counts to get percentages
2. Multiply each percentage by 100.000 (to get only numbers > 1)
3. Apply log-transformation to each percentage
4. Calculate signature score by summing up log-transformed percentages of each gene in the signature

```{r}


#Multiply all values by 100000 before taking the log, in order not to have any value below 1 (and thus no log-value below 0)
log100000 <- function(x) {
  y <- log(x*100000)
  return(y)
}

#initialize tables
score = NULL
score$cells <- all.myeloid@assays[["RNA"]]@data@Dimnames[[2]]
attr.score <- as.data.frame(score)
supp.score <- as.data.frame(score)

#calculate values for each feature and add to tables
for (i in T_attr) {
    percent.i <- GetAssayData(object=all.myeloid[["RNA"]], slot = 'counts')[i,] / Matrix::colSums(x = GetAssayData(object = all.myeloid[["RNA"]], slot = 'counts'))
    logpercent.i <- sapply(percent.i, FUN = log100000)
    logpercent.i <- ifelse(logpercent.i == "-Inf" | logpercent.i == "Inf", 0, logpercent.i)
    logpercent.i <- as.data.table(logpercent.i)
    colnames(logpercent.i) <- i
    attr.score <- cbind(attr.score, logpercent.i)
}

for (i in T_supp) {
    percent.i <- GetAssayData(object=all.myeloid[["RNA"]], slot = 'counts')[i,] / Matrix::colSums(x = GetAssayData(object = all.myeloid[["RNA"]], slot = 'counts'))
    logpercent.i <- sapply(percent.i, FUN = log100000)
    logpercent.i <- ifelse(logpercent.i == "-Inf" | logpercent.i == "Inf", 0, logpercent.i)
    logpercent.i <- as.data.table(logpercent.i)
    colnames(logpercent.i) <- i
    supp.score <- cbind(supp.score, logpercent.i)
}

#sum up all values per cell and attach to Seurat object metadata
rownames(attr.score) <- attr.score$cells
attr.score$cells <- NULL
attr.score$sum <- rowSums(attr.score)
all.myeloid$attr <- attr.score$sum

rownames(supp.score) <- supp.score$cells
supp.score$cells <- NULL
supp.score$sum <- rowSums(supp.score)
all.myeloid$supp <- supp.score$sum
```

### Calculate contribution of each gene to the T-cell-associated signatures after log-normalization

```{r}
# calculate contribution of each gene to the log-normalized attr/supp scores
logGeneContribution <- function(object, gene) {
  contrib.cell <- object[,gene]/object[,"sum"]
  contrib.cell <- ifelse(contrib.cell == "-Inf", 0, contrib.cell)
  contr.gene <- mean(contrib.cell, na.rm = TRUE)
  return(contr.gene)
}

supp_contr <- sapply(T_supp, logGeneContribution, object=supp.score)
names(supp_contr) <- T_supp
head(supp_contr)

attr_contr <- sapply(T_attr, logGeneContribution,object=attr.score)
names(attr_contr) <- T_attr
head(attr_contr)


## Plot
supp.df <- data.frame(gene = T_supp, contribution = supp_contr)
p_contr_supp = ggplot(supp.df, aes(x=gene, y=contribution))+
  geom_col()+
  theme(panel.background = element_blank())+
  coord_flip()

attr.df <- data.frame(gene = T_attr, contribution = attr_contr)
p_contr_attr = ggplot(attr.df, aes(x=gene, y=contribution))+
  geom_col()+
  theme(panel.background = element_blank())+
  coord_flip()

pdf(paste0(out.path, "Fig4S_M1M2_gene_contribution.pdf"), width = 10, height = 15)
p <- p_contr_supp / p_contr_attr
p
dev.off()
```

## Single cell plots

```{r}
attr_supp_sc <- all.myeloid[[c("attr", "supp", "TIG", "orig.ident", "cluster", "metacluster")]]

#Randomize rows for plotting
attr_supp_sc_random <- attr_supp_sc[sample(nrow(attr_supp_sc)),]

## Single Cell Scatterplot
p = ggplot(attr_supp_sc_random, aes(x=attr, y=supp)) +
  geom_point(aes(color=TIG), alpha = 0.5, size = 1)+
  geom_smooth(method = lm, se =TRUE, color = "black")+
  stat_cor(method="spearman")+
  theme(axis.line.x = element_line(colour = "black", size = 0.25),
        axis.line.y = element_line(colour = "black", size = 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.background = element_blank())


## Cluster-based Boxplot
colnames(attr_supp_sc) <- c("attr", "supp", "TIG", "sample", "cluster", "metacluster")
attr_supp_sc_gat <- gather(attr_supp_sc, "attr", "supp", key= "signature", value="signature_value")

p = ggplot(attr_supp_sc_gat, aes(x=cluster, y=signature_value))+
  geom_boxplot( width=0.8, aes(fill=metacluster))+
  facet_wrap(~signature, scales = "free")+
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90, hjust = 1))

pdf(paste0(out.path, "Fig4S_attrsupp_scBoxplots_cluster.pdf"), width = 10, height = 8)
p
dev.off()

## TIG comparison
p1 = ggplot(attr_supp_sc_gat, aes(x=signature, y =signature_value, fill = TIG))+
  facet_wrap(~signature, scales = "free")+
  geom_violin(aes(fill=TIG))+
  geom_boxplot(position=position_dodge(width=0.9), width=0.1, color="black", aes(fill=TIG))+
  stat_compare_means(method = "wilcox.test", aes(label=..p.signif..))+
  theme(axis.line.x = element_line(colour = "black", size = 0.25),
        axis.line.y = element_line(colour = "black", size = 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_blank())
```

## Cluster means

```{r}
# Get mean attr and supp percentages per cluster
Idents(all.myeloid) <- all.myeloid$cluster
attrsupp_means <- data.frame(attr_signature = double(),
                              supp_signature = double())
clusterlist <- levels(all.myeloid$cluster)

for (i in clusterlist) {
  cluster.i <- subset(x= all.myeloid, idents = i)
  newline <- c(mean(cluster.i$attr), mean(cluster.i$supp))
  attrsupp_means <- rbind(attrsupp_means, newline)
  rm(cluster.i)
}

attrsupp_means$cluster <- clusterlist
colnames(attrsupp_means) <- c("attr_signature", "supp_signature", "cluster")
#write.csv(attrsupp_means, file = paste0(out.path, "cluster_means_excl_cl9.csv"), row.names=FALSE)

##add metacluster ids
attrsupp_means$metacluster <- c(rep("Monocyte",2), rep("TAM",7), rep("DC",5), "M-proliferating")

#Scatterplot of cluster means
p = ggplot(attrsupp_means, aes(x=attr_signature, y=supp_signature)) +
  geom_point(aes(color=metacluster), size=3)+
  geom_text_repel(label=attrsupp_means$cluster)+
  geom_smooth(method = lm, color = "black", se=FALSE)+
  stat_cor(method="spearman")+
  theme(axis.line.x = element_line(colour = "black", size = 0.25),
        axis.line.y = element_line(colour = "black", size = 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.background = element_blank())+
  xlab("Random 15 gene signature")

```

## Sample means

```{r}
# Get mean attr and supp percentages per patient
Idents(all.myeloid) <- all.myeloid$orig.ident
attrsupp_means <- data.frame(attr_signature = double(),
                              supp_signature = double())
patientlist <- levels(all.myeloid$orig.ident)

for (i in patientlist) {
  patient.i <- subset(x= all.myeloid, idents = i)
  newline <- c(mean(patient.i$attr), mean(patient.i$supp))
  attrsupp_means <- rbind(attrsupp_means, newline)
  rm(patient.i)
}

attrsupp_means$patient <- patientlist
colnames(attrsupp_means) <- c("attr_signature", "supp_signature", "patient")
#write.csv(attrsupp_means, file = paste0(out.path, "patient_means_excl_cl9.csv"), row.names=FALSE)

##add TIGs
attrsupp_means$TIG <- c(rep("TIG2", 7), rep("TIG3", 7))

#Scatterplot of cluster means
p = ggplot(attrsupp_means, aes(x=attr_signature, y=supp_signature)) +
  geom_point(aes(color=TIG), size=4)+
  #geom_text_repel(label=attrsupp_means$patient)+
  geom_smooth(method = lm, color = "black", se=FALSE)+
  stat_cor(method="spearman")+
  theme(axis.line.x = element_line(colour = "black", size = 0.25),
        axis.line.y = element_line(colour = "black", size = 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.background = element_blank())

## TIG comparison (patient-level)
attrsupp_sample_gat <- gather(attrsupp_means, "attr_signature", "supp_signature", key= "signature", value="signature_value")

p = ggplot(attrsupp_sample_gat, aes(x=TIG, y =signature_value, fill = TIG))+
  geom_boxplot(aes(fill=TIG))+
  geom_point(position=position_dodge(width=0.75), aes(group=TIG))+
  facet_wrap(~signature)+
  stat_compare_means(method = "wilcox.test", aes(label=..p.format..))+
  theme(axis.line.x = element_line(colour = "black", size = 0.25),
        axis.line.y = element_line(colour = "black", size = 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.background = element_blank(),
        strip.background = element_blank())
```

### Calculate M1/M2 signatures using log-transformation of percentages for each gene
To avoid dominance of few genes when calculating the signature score: Log-normalization

1. For each cell, divide counts of gene i through total counts to get percentages
2. Multiply each percentage by 100.000 (to get only numbers > 1)
3. Apply log-transformation to each percentage
4. Calculate signature score by summing up log-transformed percentages of each gene in the signature

```{r}


#Multiply all values by 100000 before taking the log, in order not to have any value below 1 (and thus no log-value below 0)
log100000 <- function(x) {
  y <- log(x*100000)
  return(y)
}

#initialize tables
score = NULL
score$cells <- all.myeloid@assays[["RNA"]]@data@Dimnames[[2]]
M1.score <- as.data.frame(score)
M2.score <- as.data.frame(score)

#calculate values for each feature and add to tables
for (i in M1_features) {
    percent.i <- GetAssayData(object=all.myeloid[["RNA"]], slot = 'counts')[i,] / Matrix::colSums(x = GetAssayData(object = all.myeloid[["RNA"]], slot = 'counts'))
    logpercent.i <- sapply(percent.i, FUN = log100000)
    logpercent.i <- ifelse(logpercent.i == "-Inf" | logpercent.i == "Inf", 0, logpercent.i)
    logpercent.i <- as.data.table(logpercent.i)
    colnames(logpercent.i) <- i
    M1.score <- cbind(M1.score, logpercent.i)
}

for (i in M2_features) {
    percent.i <- GetAssayData(object=all.myeloid[["RNA"]], slot = 'counts')[i,] / Matrix::colSums(x = GetAssayData(object = all.myeloid[["RNA"]], slot = 'counts'))
    logpercent.i <- sapply(percent.i, FUN = log100000)
    logpercent.i <- ifelse(logpercent.i == "-Inf" | logpercent.i == "Inf", 0, logpercent.i)
    logpercent.i <- as.data.table(logpercent.i)
    colnames(logpercent.i) <- i
    M2.score <- cbind(M2.score, logpercent.i)
}

#sum up all values per cell and attach to Seurat object metadata
rownames(M1.score) <- M1.score$cells
M1.score$cells <- NULL
M1.score$sum <- rowSums(M1.score)
all.myeloid$M1 <- M1.score$sum

rownames(M2.score) <- M2.score$cells
M2.score$cells <- NULL
M2.score$sum <- rowSums(M2.score)
all.myeloid$M2 <- M2.score$sum
```

## Compare TIGs (sample means) for all 4 signatures

```{r}
# Get mean scores per patient
Idents(all.myeloid) <- all.myeloid$orig.ident
all_scores <- data.frame(attr_signature = double(),
                      supp_signature = double(),
                      M1_score = double(),
                      M2_score = double())
patientlist <- levels(all.myeloid$orig.ident)

for (i in patientlist) {
  patient.i <- subset(x= all.myeloid, idents = i)
  newline <- c(mean(patient.i$attr), mean(patient.i$supp), mean(patient.i$M1), mean(patient.i$M2))
  all_scores <- rbind(all_scores, newline)
  rm(patient.i)
}

all_scores$patient <- patientlist
colnames(all_scores) <- c("attr", "supp", "M1", "M2", "patient")

#add TIGs
all_scores$TIG <- c(rep("TIG2", 7), rep("TIG3", 7))

## TIG comparison (patient-level)
all_scores_gat <- gather(all_scores, "attr", "supp", "M1", "M2", key= "signature", value="signature_value")

p = ggplot(all_scores_gat, aes(x=TIG, y =signature_value, fill = TIG))+
  geom_boxplot(aes(fill=TIG))+
  geom_point(position=position_dodge(width=0.75), aes(group=TIG))+
  facet_wrap(~signature, scales = "free", ncol = 4)+
  stat_compare_means(method = "wilcox.test", aes(label=..p.format..))+
  theme(axis.line.x = element_line(colour = "black", size = 0.25),
        axis.line.y = element_line(colour = "black", size = 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.background = element_blank(),
        strip.background = element_blank())
```

### Single cell scatterplots

```{r}
M1M2_sc <- all.myeloid[[c("M1", "M2", "TIG", "orig.ident", "cluster", "metacluster")]]

#Randomize rows for plotting
M1M2_sc_random <- M1M2_sc[sample(nrow(M1M2_sc)),]

## Single Cell Scatterplot
p = ggplot(M1M2_sc_random, aes(x=M1, y=M2)) +
  geom_point(aes(color=TIG), alpha = 0.5, size = 1)+
  geom_smooth(method = lm, se =TRUE, color = "black")+
  stat_cor(method="spearman")+
  theme(axis.line.x = element_line(colour = "black", size = 0.25),
        axis.line.y = element_line(colour = "black", size = 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.background = element_blank())
```





