---
title: "Fig5_allCells_CPDB"
author: "Sandra Tietscher"
date: "6/2/2020"
output: html_document
---

## Load libraries and data, tidy cell type levels

```{r}
library(ggplot2)
library(Seurat)
library(tidyverse)
library(data.table)
library(patchwork)
library(ggpubr)
library(igraph)
library(viridis)
library(tidyr)
library(igraph)

path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/"
out.path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/LR_analysis/cpdb/Fig5_allCells/"
all.merged <- readRDS(file = paste0(path, "merged_complete_inclCelltype.rds"))

# classify pDCs as myeloid cells for this large overview
ct_levels <- levels(all.merged$cell.type)
ct_levels[7] <- "mast cell/basophil"
ct_levels[9] <- "myeloid"
levels(all.merged$cell.type) <- ct_levels

## reorder cell type levels
old.levels <- levels(all.merged$cell.type)
all.merged$cell.type <- factor(all.merged$cell.type, levels =  old.levels[c(6,4,3,1,2,5,7,8)])
Idents(all.merged) <- all.merged$cell.type

#subsample to max. 10.000 cells per cell type
Idents(all.merged) <- all.merged$cell.type
all.sub <- SubsetData(object = all.merged, max.cells.per.ident=3000, random.seed=11)
```

# Prepare input files for CellPhoneDB

```{r}

### Generate COUNTS file ###
#extract raw count_raw & normalize
count_raw <- all.sub@assays$RNA@counts
count_norm <- apply(count_raw, 2, function(x) (x/sum(x))*10000)
#Save COUNTS file with HNGC IDs
write.table(count_norm, paste0(out.path, "count_subsampled_3000celltype.txt"), sep="\t", quote=F)

#generate META file
meta_data <- cbind(rownames(all.sub@meta.data), all.sub@meta.data[,"cell.type", drop=F])
write.table(meta_data, paste0(out.path, "meta_subsampled_3000celltype.txt"), sep="\t", quote=F, row.names=F)
```

## Run CellPhoneDB using Python
Run as an one-liner in the console:

cellphonedb method statistical_analysis meta_subsampled_3000celltype.txt count_subsampled_3000celltype.txt --counts-data hgnc_symbol --project-name allCells_sub3000 --result-precision 4



# Read in CellPhoneDB output

```{r}
in.path <- "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/LR_analysis/cpdb/Fig5_allCells/out/allCells_sub3000/"

sign.means <- as.matrix(read.table(file=paste0(in.path, "significant_means.txt"), sep = "\t"))
colnames(sign.means) <- sign.means[1,]
sign.means <- sign.means[-1,]
#inspect
sign.means.df <- as.data.frame(sign.means)
sign.means.df[1:10, 1:15]
```

# Count number of significant interactions for each pair & plot

```{r}
#remove columns with metadata
sign.means.nometa <- sign.means[,-(1:12)]

#count significant interactions for each (directed) pair
n.int <- colSums(sign.means.nometa != "")
#get pair names
pairs <- as.data.frame(colnames(sign.means.nometa))
pairs <- separate(pairs, col="colnames(sign.means.nometa)", into=c("A", "B"), sep="\\|", remove = TRUE)
#add number of significant interactions
pairs$interactions <- n.int


###
# CellPhoneDB does not actually give directed data --> transform into undirected list by adding counts for A->B and B->A for each pair
###

directed_g <- graph_from_data_frame(pairs, directed=TRUE)
undir_g <- as.undirected(directed_g, mode = "collapse", edge.attr.comb = "sum")
#read out the undirected edge list
res <- data.frame(get.edgelist(undir_g), get.edge.attribute(undir_g, "interactions"))
colnames(res) <- c("node1", "node2", "interactions")

#prepare undirected matrix
adj_matrix <- spread(res[,1:3], key = node2, value = interactions)
rownames(adj_matrix) <- adj_matrix[,1]
adj_matrix <- as.matrix(adj_matrix[,-1])
#transpose
adj_matrix_t <- t(adj_matrix)
# convert all NAs to 0
adj_matrix[is.na(adj_matrix)] <- 0
adj_matrix_t[is.na(adj_matrix_t)] <- 0
#Sum matrices (auto-interactions will be counted twice, but this is fair as they go in both directions)
matrix_compl <- adj_matrix + adj_matrix_t

res_compl <- as.data.frame(matrix_compl)
res_compl$A <- rownames(res_compl)
res_compl <- gather(res_compl, key = "B", value = "interactions", 1:8)
res_compl$A <- factor(res_compl$A, levels = c("epithelial", "endothelial", "fibroblast", "T/NK cell", "myeloid", "B cell", "mast cell/basophil", "plasma cell"))
res_compl$B <- factor(res_compl$B, levels = c("epithelial", "endothelial", "fibroblast", "T/NK cell", "myeloid", "B cell", "mast cell/basophil", "plasma cell"))


#plot heatmap of interaction counts
p <- ggplot(res_compl)+
  geom_tile(aes(A, B, fill = interactions))+
  scale_fill_gradientn(colours=viridis(100))+
  theme(panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=1))+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Number of significant interactions")

```


