---
title: "Fig5_TIG_comparison"
author: "Sandra Tietscher"
date: "6/3/2020"
output: html_document
---

## Load libraries and data, tidy cell type levels

```{r}
library(ggplot2)
library(Seurat)
library(tidyverse)
library(data.table)
library(patchwork)
library(ggpubr)
library(SingleCellSignalR)
library(igraph)
library(viridis)

path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/"
out.path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/figures_for_paper/" 
all.merged <- readRDS(file = paste0(path, "merged_complete_inclCelltype.rds"))

#Set wd
dir.orig <- "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/LR_analysis/SCSR/Fig5_TIG_comparison/"

# classify pDCs as myeloid cells for this large overview
ct_levels <- levels(all.merged$cell.type)
ct_levels[7] <- "mast cell/basophil"
ct_levels[9] <- "myeloid"
levels(all.merged$cell.type) <- ct_levels

## reorder cell type levels
old.levels <- levels(all.merged$cell.type)
all.merged$cell.type <- factor(all.merged$cell.type, levels =  old.levels[c(6,4,3,1,2,5,7,8)])
Idents(all.merged) <- all.merged$cell.type
```


## Detect paracrine  and autocrine interactions for each individual patient (broad cell types)


```{r}
patient.list <- c("TBB075", "TBB102", "TBB111", "TBB129", "TBB165", "TBB171", "TBB214", "TBB226", "TBB330", "TBB338", "TBB011", "TBB035", "TBB212", "TBB184")
Idents(all.merged) <- all.merged$orig.ident

LR.matrix.list <- vector(mode = "list", length = 14)
names(LR.matrix.list) <- patient.list

for (patient in patient.list) {
  sub <- subset(all.merged, idents=patient)

  dir <- (paste0(dir.orig, "all_celltypes/", patient))
  dir.create(dir)
  setwd(dir)
  
cluster = as.numeric(sub$cell.type)
#c.names = levels(sub$cell.type)
c.names = c("epithelial", "endothelial", "fibroblast", "TNK cell", "myeloid",  "Bcell",  "granulocyte", "PC")
data = data.frame(sub[["RNA"]]@data)

#################
### paracrine ###
#################

##Calculate LR scores for each cell type combination
#Set s.score to 0 in order to get the score for each LR pair that is present (whether "significant" or not)
#Tolerance for autocrine/paracrine was set to 100%
setwd(dir)
signal.para <- cell_signaling(data, genes = rownames(data), cluster, int.type = "paracrine", c.names = c.names, s.score = 0.0, logFC = NULL,  species = "homo sapiens", tol = 1, write = TRUE, verbose = TRUE)

### assemble "cell pair x LR pair" data frame ###
signal.para <- lapply(signal.para, function(x) unique(x))
signal.para <- lapply(signal.para, function(x) {rownames(x) <- paste0(x[,1], "_", x[,2]);x})
signal.para <- lapply(signal.para, function(x) {x$LR <- paste0(x[,1], "_", x[,2]);x})
signal.simple.para <- lapply(signal.para, function(x) x[,-c(1:3)])
#pull everything together in one matrix per sample (cluster_pairs x LR_pairs)
signal.simple.para <- Reduce(function(x, y) merge(x,y, by= "LR", all.x=TRUE, all.y=TRUE), signal.simple.para)
colnames(signal.simple.para) <- c("LR", names(signal.para))
signal.simple.para[is.na(signal.simple.para)] <- 0


#################
### autocrine ###
#################

#Tolerance for autocrine/paracrine was set to 0%
setwd(dir)
signal.auto <- cell_signaling(data, genes = rownames(data), cluster, int.type = "autocrine", c.names = c.names, s.score = 0.0, logFC = NULL,  species = "homo sapiens", tol = 0, write = TRUE, verbose = TRUE)

### assemble "cell pair x LR pair" data frame ###

signal.auto <- lapply(signal.auto, function(x) unique(x))
signal.auto <- lapply(signal.auto, function(x) {rownames(x) <- paste0(x[,1], "_", x[,2]);x})
signal.auto <- lapply(signal.auto, function(x) {x$LR <- paste0(x[,1], "_", x[,2]);x})
signal.simple.auto <- lapply(signal.auto, function(x) x[,-c(1:3)])
#pull everything together in one matrix per sample (cluster_pairs x LR_pairs)
signal.simple.auto <- Reduce(function(x, y) merge(x,y, by= "LR", all.x=TRUE, all.y=TRUE), signal.simple.auto)
colnames(signal.simple.auto) <- c("LR", names(signal.auto))
signal.simple.auto[is.na(signal.simple.auto)] <- 0

### subset only autocrine pairs ###
signal.simple.auto <- signal.simple.auto[,c(1,2,11, 20, 29, 38, 47, 56, 65)]


#############
### merge ###
#############

LR_full <- merge(signal.simple.para, signal.simple.auto, by = "LR", all = TRUE)
## Exclude interaction pairs that are clearly not extracellular/surface: Heat-Shock Proteins, Ubiquitin-related, RPS
exclude_list <- c("HSP", "UBB", "UBC", "UBA", "RPS", "RPL")
LR_full <- dplyr::filter(LR_full, !substr(LR, 1, 3) %in% exclude_list)

#write out as csv and add to the matrix list
write.csv(LR_full, file=paste0(dir, "/LR_clusterpair_matrix.csv"), row.names = FALSE)
LR.matrix.list[[patient]] <- LR_full

}

```

### Read the individual tables back in into list form

```{r}
list.dirs(paste0(dir.orig, "all_celltypes/"))
sample.dirs <- list.dirs(paste0(dir.orig, "all_celltypes/"))[-c(1:57, 86:93)]
odd_indexes<-seq(1,27,2)
sample.dirs <- sample.dirs[odd_indexes]
LRscore.dirs <- paste0(sample.dirs, "/LR_clusterpair_matrix.csv")
LRscore.list <- lapply(LRscore.dirs, function(x) read.csv(x, header=T))

samples <- sub("(.*?)TBB" , "TBB", sample.dirs)
TIG <- c("TIG2", "TIG3", "TIG3","TIG3", "TIG2", "TIG2","TIG2","TIG2","TIG2","TIG3","TIG3","TIG3","TIG3","TIG2")
names(LRscore.list) <- samples

cell.pair.list <- colnames(LRscore.list[[1]])[-1]
```

### Prepare separate data frames for each specific cell interaction
Subset interactions that are significantly higher in TIG2 or TIG3 (p<= 0.05) and have a mean LRscore of at least 0.4 in that TIG.

```{r}
dir <- paste0(dir.orig, "all_celltypes/")

#initialize data frames
TIG2.high <- data.frame(ligand = character(),
                        receptor = character(),
                        p = double(),
                        mean_meax = double(),
                        higher = character())
TIG3.high <- TIG2.high

for (i in cell.pair.list) {
  
  #set directory
  dirx <- paste0(dir, i, "/")
  dir.create(dirx)
  
  ## select only cell-cell column and LR names
  TM.table <- lapply(LRscore.list, function(x) x[, c("LR", i )])
  ## attach sample names and TIG to cell-cell column
  for (x in 1:14) {
    colnames(TM.table[[x]])[[2]] <- paste0("myeloid.TNK_", samples[x], "_", TIG[x])
  }

  ## merge by LR names (includes also LRs that do not appear in all samples) and check for     duplicates
  TM.table <- Reduce(function(x, y) merge(x,y, by= "LR", all.x=TRUE, all.y=TRUE), TM.table)
  #dups <- TM.table[duplicated(TM.table$LR),]
  #View(dups)
  #TM.table <- TM.table[!duplicated(TM.table$interacting_pair),]
  TM.table[is.na(TM.table)] <- 0
  #write out table
  write.csv(TM.table, file=paste0(dirx, i, "_full_matrix.csv"))

  #format the table
  TM.dat <- gather(TM.table, key="ID", value = "LR_score", 2:15)
  TM.dat <- separate(TM.dat, col = "ID", into=c("cell.pair", "sample", "TIG"), sep="_", remove=TRUE)

  
  ## Find interactions that are overrepresented in one of the TIGs

  #Calculate Wilcoxon test for each LR pair
  pvalues <- TM.dat %>% 
  group_by(LR) %>% 
  summarise(p=wilcox.test(LR_score~TIG)$p.value)

  #Calculate means for TIG2 and TIG3
  mean <- TM.dat %>%
    group_by(LR, TIG) %>%
    summarise(mean=mean(LR_score))
  max_mean <- mean %>%
    group_by(LR) %>%
    summarise(mean_max=max(mean))
  
  diff_mean <- mean %>%
    group_by(LR) %>%
    summarise(diff=diff(mean))
  diff_mean$higher <- ifelse(diff_mean$diff > 0, "TIG3", "TIG2")
  TIG_higher <- diff_mean[,-2]

  #Merge p-values and max_mean with LR table
  p_mean <- merge(pvalues, max_mean, by = "LR")
  tdat <- merge(TM.dat, p_mean, by.x = "LR", by.y ="LR", all.x = TRUE)
  tdat$p.wt <- paste0('p=',round(tdat$p, digits=4))

  #Exclude genes that have the max(mean) < 0.5. & sort remaining genes by lowest pvalue
  p_mean_sub <- subset(p_mean, p_mean$mean_max >= 0.4)
  setorder(p_mean_sub, p, na.last=TRUE)
  TM.top <- top_n(p_mean_sub, -500, p)
  write.csv(TM.top, file=paste0(dirx, i, "_top500LR.csv"))

  TM.sign.05 <- subset(tdat, tdat$p<=0.05 & tdat$mean_max>=0.4)
  TM.sign.01 <- subset(tdat, tdat$p<=0.01 & tdat$mean_max>=0.4)
  
  # GOI <- c("CD274_PDCD1", "CD80_CTLA4", "LGALS9_HAVCR2", "CD47_SIRPG", "IL15_IL2RB", "CXCL9_CXCR3", "CCL20_CXCR3", "CCL4_CCR1")
  #GOI2 <- c("CCL18_CCR8")
  GOI_TM <- c("LGALS9_HAVCR2", "CSF1_CSF1R", "FLT3LG_FLT3", "CCL3_CCR1")
  GOI_MT <- c("CD274_PDCD1", "CD80_CTLA4", "CXCL9_CXCR3", "IL15_IL2RB")
  GOI_EpT <- c("CXCL9_CXCR3", "LGALS9_HAVCR2", "CD274_PDCD1", "CCL3_CCR1")
  TM.select <- subset(tdat, LR %in% GOI_EpT)

  #plot LR pairs that are significantly different
  if (nrow(TM.sign.05) > 0) {
  p05 = ggplot(TM.sign.05, aes(x= TIG, y = LR_score, fill=TIG))+
    geom_boxplot()+
    geom_point()+
    facet_wrap(~LR+p.wt, scales="free", ncol=8, strip.position="bottom")+
    theme(axis.text.x=element_blank(), axis.ticks.x = element_blank())+
    theme(panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=1))

  ggsave(paste0(dirx, i, "_p005.png"), width = 16, height=20)
  }

  if (nrow(TM.sign.01) > 0) {
  p01 = ggplot(TM.sign.01, aes(x= TIG, y = LR_score, fill=TIG))+
    geom_boxplot()+
    geom_point()+
    facet_wrap(~LR+p.wt, scales="free", ncol=5, strip.position="bottom")+
    theme(axis.text.x=element_blank(), axis.ticks.x = element_blank())+
    theme(panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=1))
  
  ggsave(paste0(dirx, i, "_p001.png"), width = 10, height = 12)
  }
  
  ### merge p values and means to get complete TIG2vsTIG3 table
  mean <- spread(mean, key = "TIG", value = "mean")
  comp <- merge(mean, pvalues, by="LR")
  comp$delta <- comp$TIG2 - comp$TIG3
  comp$cellpair <- i
  write.csv(comp, file=paste0(dirx, i, "TIG2vsTIG3.csv"))
}
```

## Merge & plot LR pairs of interest for Ep-T-M interactions

```{r}
dir <- paste0(dir.orig, "all_celltypes/")
cell.pair.list <- cell.pair.list[c(25, 22, 32, 29, 3, 4)]
dir.list <- paste0(dir, cell.pair.list, "/", cell.pair.list, "TIG2vsTIG3.csv")

#initialize data frame and read in tables
comp.all <- data.frame(X=character(),
                        LR = character(),
                        TIG2 = double(),
                        TIG3 = double(),
                        p = double(),
                        delta = double(),
                        cellpair = character())

for (i in dir.list) {
  df <- read.csv(i)
  comp.all <- rbind(comp.all, df)
}

comp.all[,1] <- NULL
comp.all$maxmean <- pmax(comp.all$TIG2, comp.all$TIG3)
comp.all$TIG <- ifelse(comp.all$delta > 0, "TIG2", "TIG3")

#all LR pair/cluster pair combinations where the max mean is < 0.4 --> assign delta = 0 and p = n.s.
comp.all$delta <- ifelse(comp.all$maxmean <= 0.4, NA, comp.all$delta)
comp.all$p<- ifelse(comp.all$maxmean <= 0.4,  NA , comp.all$p)
comp.all$TIG<- ifelse(comp.all$maxmean <= 0.4,  NA , comp.all$TIG)

#all LR pair/cluster pair combinations where p is above 0.05 --> assign delta = 0 and p = n.s.
comp.all$delta <- ifelse(comp.all$p > 0.05,  NA, comp.all$delta)
comp.all$p<- ifelse(comp.all$p > 0.05,  NA , comp.all$p)
comp.all$TIG<- ifelse(comp.all$p > 0.05,  NA , comp.all$TIG)

#Change cell pair names
levels(comp.all$cellpair) <- c("T/NK -> M", "T/NK -> tumor", "M -> T/NK", "M -> tumor", "tumor -> T/NK", "tumor -> M")

#Reorder LR pairs
comp.all$LR <- factor(comp.all$LR, levels = rev(levels(comp.all$LR)))

GOI <- c("CD274_PDCD1", "CD80_CTLA4","CD86_CTLA4", "LGALS9_HAVCR2", "PVR_TIGIT", "FASLG_FAS", 
        "CCL3_CCR5", "CCL3_CCR1", "CCL5_CCR5", "CCL5_CCR1", "CCL4_CCR1", "CCL4_CCR5", "CXCL9_CXCR3",  
         "VEGFA_NRP1", "CSF1_CSF1R",  "IL15_IL2RB", "IL1RN_IL1R2", "IL10_IL10RA", "FLT3LG_FLT3",
         "EREG_ERBB4", "HBEGF_ERBB4", "THBS1_ITGB1", "THBS1_ITGA4", "TNFSF10_TNFRSF10B")
comp.small <- comp.all[comp.all$LR %in% GOI,]
#comp.small$LR <- factor(comp.small$LR, levels = rev(GOI))
comp.small$LR <- factor(comp.small$LR, levels = GOI)
comp.small$padj <- ifelse(comp.small$p < 0.001, 9, ifelse(comp.small$p < 0.01, 5, ifelse(comp.small$p < 0.05, 2, 0)))


p <- ggplot(comp.small)+
  geom_tile(aes(cellpair, LR, fill = TIG))+
  scale_fill_manual(name= "Upregulated in", values = c("#F8766D", "#00BFC4", "grey50"))+
  #scale_alpha(name = "\u0394 LR-score", range = c(0.4, 1))+
  geom_point(aes(cellpair, LR, size = padj), color = "black", shape = 16)+
  scale_size_area("p-values",breaks=c(2, 5, 9),labels=c("<0.05", "<0.01", "<0.001"), limits = c(0,9), max_size = 9)+
  theme(axis.text.x = element_text(angle=90, hjust = 0, vjust = 1, size=12), 
        axis.text.y = element_text(size=12),
        axis.title.x =element_text(size=15),
        axis.title.y =element_text(size=15),
        panel.background= element_blank(), panel.border = element_rect(color="black", fill=NA, size=2))+
  labs(x="Cell type pair", y="Interacting partners")+
  scale_x_discrete(position = "top")+
  coord_flip()
 


```