---
title: "Fig5_SCSR_allCells"
author: "Sandra Tietscher"
date: "5/29/2020"
output: html_document
---

## Load libraries and data, tidy cell type levels

```{r}
library(ggplot2)
library(Seurat)
library(tidyverse)
library(data.table)
library(patchwork)
library(ggpubr)
library(SingleCellSignalR)
library(igraph)
library(viridis)

path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/"
out.path = "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/figures_for_paper/" 
all.merged <- readRDS(file = paste0(path, "merged_complete_inclCelltype.rds"))

#Set wd
dir <- "~/Code/10x/10x_Run1_Run2_Pilot_comb/output/LR_analysis/SCSR/Fig5_allCells/"

# classify pDCs as myeloid cells for this large overview
ct_levels <- levels(all.merged$cell.type)
ct_levels[7] <- "mast cell/basophil"
ct_levels[9] <- "myeloid"
levels(all.merged$cell.type) <- ct_levels

## reorder cell type levels
old.levels <- levels(all.merged$cell.type)
all.merged$cell.type <- factor(all.merged$cell.type, levels =  old.levels[c(6,4,3,1,2,5,7,8)])
Idents(all.merged) <- all.merged$cell.type
```

#Run SCSR on main cell types (for paracrine and autocrine interactions)

```{r}
#subsample to max. 10.000 cells per cell type
Idents(all.merged) <- all.merged$cell.type
all.sub <- SubsetData(object = all.merged, max.cells.per.ident=10000, random.seed=11)

#prepare data
cluster = as.numeric(all.sub$cell.type)
#c.names = levels(sub$cell.type)
levels(all.sub$cell.type)
c.names <- c("epithelial", "endothelial", "fibroblast", "Tcell", "myeloid", "Bcell", "mast_basophil", "plasma")
data = data.frame(all.sub[["RNA"]]@data)


#################
### paracrine ###
#################

##Calculate LR scores for each cell type combination
#Set s.score to 0 in order to get the score for each LR pair that is present (whether "significant" or not)
#Tolerance for autocrine/paracrine was set to 100%
setwd(dir)
signal.para <- cell_signaling(data, genes = rownames(data), cluster, int.type = "paracrine", c.names = c.names, s.score = 0.0, logFC = NULL,  species = "homo sapiens", tol = 1, write = TRUE, verbose = TRUE)

### assemble "cell pair x LR pair" data frame ###
signal.para <- lapply(signal.para, function(x) unique(x))
signal.para <- lapply(signal.para, function(x) {rownames(x) <- paste0(x[,1], "_", x[,2]);x})
signal.para <- lapply(signal.para, function(x) {x$LR <- paste0(x[,1], "_", x[,2]);x})
signal.simple <- lapply(signal.para, function(x) x[,-c(1:3)])
#pull everything together in one matrix per sample (cluster_pairs x LR_pairs)
signal.simple <- Reduce(function(x, y) merge(x,y, by= "LR", all.x=TRUE, all.y=TRUE), signal.simple)
colnames(signal.simple) <- c("LR", names(signal.para))
signal.simple[is.na(signal.simple)] <- 0

#write out as csv
write.csv(signal.simple, file=paste0(dir, "/full_LR_matrix_paracrine.csv"), row.names = FALSE)

#################
### autocrine ###
#################

#Tolerance for autocrine/paracrine was set to 0%
setwd(dir)
signal.auto <- cell_signaling(data, genes = rownames(data), cluster, int.type = "autocrine", c.names = c.names, s.score = 0.0, logFC = NULL,  species = "homo sapiens", tol = 0, write = TRUE, verbose = TRUE)

### assemble "cell pair x LR pair" data frame ###

signal.auto <- lapply(signal.auto, function(x) unique(x))
signal.auto <- lapply(signal.auto, function(x) {rownames(x) <- paste0(x[,1], "_", x[,2]);x})
signal.auto <- lapply(signal.auto, function(x) {x$LR <- paste0(x[,1], "_", x[,2]);x})
signal.simple <- lapply(signal.auto, function(x) x[,-c(1:3)])
#pull everything together in one matrix per sample (cluster_pairs x LR_pairs)
signal.simple <- Reduce(function(x, y) merge(x,y, by= "LR", all.x=TRUE, all.y=TRUE), signal.simple)
colnames(signal.simple) <- c("LR", names(signal.auto))
signal.simple[is.na(signal.simple)] <- 0

### subset only autocrine pairs ###
signal.simple <- signal.simple[,c(1,2,11, 20, 29, 38, 47, 56, 65)]

#write out as csv
write.csv(signal.simple, file=paste0(dir, "/full_LR_matrix_autocrine.csv"), row.names = FALSE)
```

## Read data back in and merge autocrine and paracrine

```{r}
LR_paracrine <- read.csv(file=paste0(dir, "/full_LR_matrix_paracrine.csv"))
LR_autocrine <- read.csv(file=paste0(dir, "/full_LR_matrix_autocrine.csv"))
LR_full <- merge(LR_paracrine, LR_autocrine, by = "LR", all = TRUE)

## Exclude interaction pairs that are clearly not extracellular/surface: Heat-Shock Proteins, Ubiquitin-related, RPS
exclude_list <- c("HSP", "UBB", "UBC", "UBA", "RPS", "RPL")
LR_full <- dplyr::filter(LR_full, !substr(LR, 1, 3) %in% exclude_list)

rownames(LR_full) <- LR_full[,1]
LR_full <- LR_full[,-1]
LR_full[is.na(LR_full)] <- 0
```

## Write out Top100 interactions for each cell type pair

```{r}
for (i in colnames(LR_full)) {
  sub <- LR_full[,i]
  names(sub) <- rownames(LR_full)
  sub <- sort(sub, decreasing = TRUE)
  sub <- sub[1:100]
  write.csv(sub, file=paste0(dir, "/Top100_LR/", i, ".csv"), row.names = TRUE)
}
```

# Prepare weighted edge list by counting number of significant interactions for each pair

```{r}

#count significant interactions for each (directed) pair
n.int <- colSums(LR_full >= 0.4)
#get pair names
pairs <- as.data.frame(colnames(LR_full))
pairs <- separate(pairs, col="colnames(LR_full)", into=c("A", "B"), sep="\\.", remove = TRUE)
#add number of significant interactions
pairs$interactions <- n.int
pairs$A <- factor(pairs$A, levels = pairs$A[57:64])
pairs$B <- factor(pairs$B, levels = pairs$A[57:64])

#plot distribution of interactions counts
p <- ggplot(pairs, aes(interactions))+
  geom_density()

#plot heatmap of interaction counts
p <- ggplot(pairs)+
  geom_tile(aes(A, B, fill = interactions))+
  scale_fill_gradientn(colours=viridis(100))+
  theme(panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=1))+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Number of interactions with LRscore >= 0.4")+
  xlab("Ligand-expressing cell")+
  ylab("Receptor-expressing cell")
```

# Plot social interaction networks using igraph

```{r}
## prepare directed graph

directed_g <- graph_from_data_frame(pairs, directed=TRUE)
#set cell number as vertex attribute
cell.nr <- table(all.merged$cell.type)
directed_g<- set_vertex_attr(directed_g, "cell_nr", value = cell.nr)
get.vertex.attribute(directed_g)

## set edge color based on number of interactions
E.weight <- E(directed_g)$interactions
# Set minimum to 0 and produce viridis palette with right number of colors
E.weight <- E.weight - (min(E.weight)-1)
colors <- viridis(ceiling(max(E.weight)))
E.weight <- ceiling(E.weight)
E.colors <- unlist(sapply(E.weight, function(x) colors[x], USE.NAMES = FALSE))
directed_g<- set_edge_attr(directed_g, "color", value = E.colors) 

## Get UMAP coordinates
#load saved centroid coordinates
umap_coord <- read.csv(paste0(path, "LR_analysis/SCSR/TIG2vsTIG3/all_celltypes/umap_celltype_coord.csv"))
#order and remove pDC
umap_coord <- umap_coord[c(6,4,3,1,2,5,7,8),]
rownames(umap_coord) <- umap_coord$cluster
umap_coord <- umap_coord[,-1]
umap_coord <- as.matrix(umap_coord)

#plot
plot.igraph(directed_g, vertex.size = V(directed_g)$cell_nr/500, color=E(directed_g)$color, edge.curved=TRUE, layout=umap_coord, edge.arrow.size = 0.5)
#plot with equally sized vertices
plot.igraph(directed_g, color=E(directed_g)$color, vertex.size = 3, edge.curved=TRUE, layout=umap_coord, edge.arrow.size = 0.5)
```

### Subset most abundant cell types for graph network

```{r}
## Get UMAP coordinates
#load saved centroid coordinates
umap_coord <- read.csv(paste0(path, "LR_analysis/SCSR/TIG2vsTIG3/all_celltypes/umap_celltype_coord.csv"))
umap_coord_small <- umap_coord[c(6,4,3,1,2,5),]
umap_coord_small <- umap_coord_small[,-1]
umap_coord_small <- as.matrix(umap_coord_small)

#remove pairs involving B cell, granulocytes and plasma cells
small.celltypes <- "Bcell|plasma|mast_basophil"
LR_small <- pairs[grep(small.celltypes, pairs$A, invert=TRUE),]
LR_small <- LR_small[grep(small.celltypes, LR_small$B, invert=TRUE),]

## prepare directed graph
directed_g <- graph_from_data_frame(LR_small, directed=TRUE)

#set cell number as vertex attribute
cell.nr.small <- cell.nr[1:5]
directed_g<- set_vertex_attr(directed_g, "cell_nr", value = cell.nr.small)

## set edge color based on number of interactions
E.weight <- E(directed_g)$interactions
# Set minimum to 0 and produce viridis palette with right number of colors
E.weight <- E.weight - (min(E.weight)-1)
colors <- viridis(ceiling(max(E.weight)))
E.weight <- ceiling(E.weight)
E.colors <- unlist(sapply(E.weight, function(x) colors[x], USE.NAMES = FALSE))
directed_g<- set_edge_attr(directed_g, "color", value = E.colors) 

#plot
plot.igraph(directed_g, color=E(directed_g)$color, edge.curved=TRUE, edge.arrow.size = 0.5, layout = umap_coord_small, vertex.size=V(directed_g)$cell_nr/500, edge.width=2)

#plot reduced heatmap of interaction counts
p <- ggplot(LR_small)+
  geom_tile(aes(A, B, fill = interactions))+
  scale_fill_gradientn(colours=viridis(100))+
  theme(panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=1))+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  ggtitle("Number of interactions with LRscore >= 0.4")+
  xlab("Ligand-expressing cell")+
  ylab("Receptor-expressing cell")
```

## Find interactions that are most variable between cluster pairs

```{r}
#Start with full LR list, exclude all interactions that include plasma cells, B cells and mast_basophils
small.celltypes <- "Bcell|plasma|mast_basophil"
LR_small <- LR_full[,grep(small.celltypes, colnames(LR_full), invert=TRUE)]

################################
### Find total most variable ###
################################

#multiply everything by 1000 to get numbers > 1
LR_small.adj <- sapply(LR_small, function(x) x*1000)
#calculate coefficient of variation for each LR pair
cv=apply(LR_small.adj,1,sd)/apply(LR_small.adj,1,mean)

#sort the original matrix by variation coefficient & kick out every LR pair that scores <0.5 for all cluster pairs
LR_small.adj = LR_small[order(cv,decreasing = TRUE),]
LR_small.adj = LR_small.adj[apply(LR_small.adj,1, max)>0.5,]

n = nrow(LR_small.adj)
pheatmap::pheatmap(LR_small.adj[1:100,], cluster_cols = TRUE)


##############################################################
### Find most variable among Top100 of each cell type pair ###
##############################################################

topCV <- function(df, pair, top) {
  
  #get 100 highest-scoring interactions of the pair
  df_sub <- df[,pair]
  names(df_sub) <- rownames(df)
  df_sub <- sort(df_sub, decreasing = TRUE)
  df_sub <- df_sub[1:100]
  
  #subset whole df based on this interactions and calculate CV for each
  df_sub_LR <- df[names(df_sub),]
  df_sub_LR.adj <- sapply(df_sub_LR, function(x) x*1000)
  cv=apply(df_sub_LR.adj,1,sd)/apply(df_sub_LR.adj,1,mean)
  names(cv) <- names(df_sub)
  
  #Sort interactions based on CV and write out the Top10
  cv <- sort(cv, decreasing = TRUE)
  return(names(cv[1:top]))
}

#Apply function to every cell type pair
LR_topCV <- character()
for (i in colnames(LR_small)) {
  LR.i <- topCV(LR_small, i, top = 5)
  LR_topCV <- c(LR_topCV, LR.i)
}

#Remove duplicates
LR_topCV <- unique(LR_topCV)

#Reorder columns & plot
LR_small <- LR_small[,c(21, 1:5, 22, 6:10, 23, 11:15, 24, 16:20, 25)]
pheatmap::pheatmap(LR_small[LR_topCV,], cluster_cols = FALSE, color = viridis(100))

#################################################
### Plot only selected LR and cell type pairs ###
#################################################

LR_selected <- c("COL4A1_ITGA1", "COL6A3_ITGA1", "COL5A2_ITGA1", "COL6A1_ITGA2", "COL4A1_ITGB8", "COL1A2_ITGA11", "PDGFB_PDGFRB", "PDGFA_PDGFRB",
                 "PECAM1_CD38", "CD34_SELL", "CD34_SELP",
                 "CDH1_ITGAE", "PIP_CD4",
                 "IFNG_IFNGR2", "CCL5_CCR4",
                 "CXCL16_CXCR6", "CXCL10_CXCR3", "IL10_IL10RA")
                 
                 
                 
                 
                 

#Reorder columns & plot
LR_small <- LR_small[,c(21, 1:5, 22, 6:10, 23, 11:15, 24, 16:20, 25)]
pheatmap::pheatmap(LR_small[LR_selected,], cluster_cols = FALSE, cluster_rows = FALSE, color = viridis(100))
```